---
title: "Jiankun_Dong_HW1"
author: "Jiankun (Bob) Dong"
date: "2023-09-18"
output: html_document
---
Using R Markdown to generate the file, therefore the r code are in-line.

Problem 1: Using the given dataset\

```{r setup, include=TRUE, echo = TRUE}
#load kable lib for generating tables
library(kableExtra)
#loading the file
setwd("C:/BU/CSSE/CS555/HW1")
DaysRAW <- read.csv("./A01.csv",header = TRUE)
```
1) Loaded the csv file as DaysRaw.\
2) Drawing the histogram based on the loaded data. \

```{r , include=TRUE, echo = TRUE}
DaysPlot <- hist(DaysRAW$Days, main = "Duration of Hospital Stays", 
                 xlab = "Duration (Days)", 
                 breaks = seq(min(DaysRAW),max(DaysRAW),1), 
                 xlim = c(min(DaysRAW),max(DaysRAW)+1),
                 ylim = (c(0,25)), plot = TRUE, right = F)
```
```{r , include=FALSE, echo = FALSE}
daysMean <- mean(DaysRAW$Days)
daysMedian <- median(DaysRAW$Days)
daysVar <- var(DaysRAW$Days)
daysSD <- sd(DaysRAW$Days)
daysMin <- min(DaysRAW$Days)
daysMax <- max(DaysRAW$Days)

iqr <- quantile(DaysRAW$Days, .75)[[1]] - quantile(DaysRAW$Days,.25)[[1]]
lowerBound <- quantile(DaysRAW$Days,.25)[[1]]-iqr*1.5
upperBound <- quantile(DaysRAW$Days,.75)[[1]]+iqr*1.5
outlier <-DaysRAW$Days<lowerBound | DaysRAW$Days>upperBound
outlierDays <- DaysRAW$Days[outlier]
```
Shape: \ 
The data is right skewed.\
Center: \ 
The center of the data is `r quantile(DaysRAW$Days, .5)` days.\
Spread: \ 
The first qualtile of the data is `r quantile(DaysRAW$Days,.25)` days, and the third quantile is `r quantile(DaysRAW$Days, .75)` days. With a standard diviation of `r daysSD`.\
Outliers: \
Because the iqr of the data is `r iqr` day, we get the lower bound `r lowerBound` and upper bound `r upperBound`\
The outliers are: `r outlierDays`\

3)\
```{r include=TRUE, echo=TRUE}
daysFrame <- data.frame(
  Mean = daysMean,
  Median = daysMedian,
  SD = daysSD,
  First_Quantile = quantile(DaysRAW$Days,.25)[[1]],
  Third_Quantile = quantile(DaysRAW$Days,.75)[[1]],
  Min = daysMin,
  Max = daysMax
)
daysTable <- kable(daysFrame,"simple")
```
`r daysTable`
Because the outliers are all beyond the third quantile, and the histogram is right skewed, the best value to summarize the center of this distribution is the median `r daysMedian` days.\

The best number to describe the spread of the data is the standard deviation `r daysSD`.

Problem 2: \
4) \
part a:\

```{r include=TRUE, echo=TRUE}
# Question 2 --------------------------------------------------------------
LessThanTen <- (pnorm(10,5,3) - pnorm(0,5,3))/pnorm(0,5,3,lower.tail = FALSE)
```
The percentage of the patients in hospital for less than 10 days is `r 100*LessThanTen`%.\

part b:\

```{r include=TRUE, echo=TRUE}
n <- 35
SE <- 3/sqrt(n)
MoreThanSix <- pnorm(6,5,SE,lower.tail = FALSE)/(1-pnorm(0,5,SE))
```
Because the sample size `r n` is larger than 30, we can use the CLT.
The probability of the average of the data set being more than 6 days is `r MoreThanSix`.
