---
title: "Jiankun_Dong_CS555_HW2"
author: "Jiankun (Bob) Dong CM3226"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

Problem 1:

```{r include=TRUE}
kid_cal <- read.csv("kid_cal.csv")
Kid_meal <- kid_cal$Calories[kid_cal$trt == T]
Kid_no_meal <- kid_cal$Calories[kid_cal$trt == F]

IQparticipant <- 1.5*(quantile(Kid_meal,.75)[[1]]-quantile(Kid_meal,.25)[[1]])
outlierMeal <- Kid_meal < quantile(Kid_meal,.25)[[1]]-IQparticipant | Kid_meal > quantile(Kid_meal, .75)[[1]]+IQparticipant
#sum(outlierMeal) There are no outliers
mealFrame <- data.frame(
  Mean = mean(Kid_meal),
  Median = median(Kid_meal),
  SD = sd(Kid_meal),
  First_Quantile = quantile(Kid_meal,.25)[[1]],
  Third_Quantile = quantile(Kid_meal,.75)[[1]],
  Min = min(Kid_meal),
  Max = max(Kid_meal),
  outlier = "NULL"
)

IQNonparticipant <- 1.5*(quantile(Kid_no_meal,.75)[[1]]-quantile(Kid_no_meal,.25)[[1]])
outlierNoMeal <- Kid_no_meal< quantile(Kid_no_meal,.25)[[1]]-IQNonparticipant | Kid_no_meal > quantile(Kid_no_meal,.75)[[1]]+IQNonparticipant

noMealFrame <- data.frame(
  Mean = mean(Kid_no_meal),
  Median = median(Kid_no_meal),
  SD = sd(Kid_no_meal),
  First_Quantile = quantile(Kid_no_meal,.25)[[1]],
  Third_Quantile = quantile(Kid_no_meal,.75)[[1]],
  Min = min(Kid_no_meal),
  Max = max(Kid_no_meal),
  outliner = Kid_no_meal[outlierNoMeal]
)

```

Summary of Calorie for Participants of Meal Preparation:

```{r include=TRUE}
(mealTable <- kable(mealFrame,"simple"))
```

Summary of Calorie for Non-Participants of Meal Preparation:

```{r include=TRUE}
(nomealTable <- kable(noMealFrame,"simple"))
```

```{r include=TRUE}
par(mfrow = c(1,2))
hist(Kid_meal, main = "Calorie Distribution for Participants", 
     xlab = "Calorie", ylab = "Child Count",breaks = 10, xlim = c(100,700))
hist(Kid_no_meal, main = "Calorie Distribution for Non-Participants", 
     xlab = "Calorie", ylab = "Child Count",breaks = 10, xlim=c(100,700))
```

The graph of non-participant's meal calorie distribution follows roughly a normal distribution, with ONE outlier on the right side of the graph.

The graph of participant's meal calorie distribution also roughly follows a normal distribution, but without any outlier.

Both graph have similar shape, but non-participant's graph has a wider range of calorie.

Problem 2:\

```{r}
alpha <- 0.05
n <- length(Kid_meal)
```

Step 1: $H_0$ : $\mu 0 = 425$ $H_1$ : $\mu 1 \neq 425$ $\alpha = 0.05$ and $n=25$, df = 24\
Step 2: because the population $\sigma$ is unknown and the sample is small, use the t test where $t=\frac{\overline{x}-\mu}{\frac {S}{\sqrt{n}} }$ \
Step 3: Decision rule: Reject $H_0$ if $t \ge 2.064$ or  $t \le -2.064$\
Step 4:
```{r}
xbar <- mean(Kid_meal)
S <- sd(Kid_meal)
t <- (xbar - 425)*sqrt(25)/S
```
the t value is `r t` \
Step 5: Do not reject $H_0$ \
We do not have strong evidence that at confidence level $\alpha=0.05$ that the mean calorie consumption for those who participated in the meal preparation differ from 425.\

Problem 3:\
```{r}
meal_test <- t.test(Kid_meal,conf.level = .9)
```
The 90% confidence interval's lower bound is `r meal_test$conf.int[1]`, and upper bound is `r meal_test$conf.int[2]`. \
This means that at 90% confidence level, we will reject hypothesis that states the mean of the participant's meal calorie is below`r meal_test$conf.int[1]`, or higher than `r meal_test$conf.int[2]`. \

Problem 4:\
Step1: $H_0 : \mu1=\mu2, H1 : \mu1 \ge \mu2, \alpha=0.05, df = 21$ where non-participant correspond to $\mu2$ \
Step2: $t=\frac{(\overline x_1 - \overline x_2)-(\mu1-\mu2)}{\sqrt{\frac{S_1^2}{n_1} +\frac{S_2^2}{n_2}}}$\
Step3: Decision rule: we reject $H_0$ if $t \ge 1.721$\
Step4: 
```{r}
t_5 <- (mean(Kid_meal)-mean(Kid_no_meal))/sqrt(sd(Kid_meal)^2/25+sd(Kid_no_meal)^2/22)
```

Step5: Fail to reject $H_0$ because `r t_5` is not greater than 1.721 \
We do not have significant evidence that at $\alpha=0.05$ level that participants consume more calorie than non-participants.\

Problem 5:\
The is indeed one outlier in the non-participant dataset. Furthermore, we don't know the method the data was taken, therefore we can't say with confidence that the samples are independent.\