---
title: "Jiankun_Dong_CS555_HW3"
author: "Jiankun (Bob) Dong CM3226"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup = T}
rm(list = ls())
library("ggplot2")
setwd("C:/BU/CSSE/CS555/HW3")
a03 <- read.csv("./A03.csv")
summary(a03)
attach(a03)
```

Problem 1:\
Number of meals with fish should be on the x-axis and total mercury level should be on the y-axis. Because eating meals with fish happens before potential change in mercury level in body.\

```{r}
plot(Number.of.meals.with.fish, Total.Mercury.in.mg.g, xlab = "Number of Meal With Fish", 
     ylab = "Total Mercury Level(mg)", title("Plot of Numebr of meal with fish vs Total mercury level"), pch = 'x')
meal_bar <- mean(Number.of.meals.with.fish)
meal_sd <- sd(Number.of.meals.with.fish)
merc_bar <- mean(Total.Mercury.in.mg.g)
merc_sd <-  sd(Total.Mercury.in.mg.g)
#First we need to check for outliers before using cor function for correlation
IQR_meal <- quantile(Number.of.meals.with.fish,.75)-quantile(Number.of.meals.with.fish,.25)
meal_outlier <- Number.of.meals.with.fish < quantile(Number.of.meals.with.fish,.25)-1.5*IQR_meal | Number.of.meals.with.fish > quantile(Number.of.meals.with.fish,.75)+1.5*IQR_meal
IQR_merc <- quantile(Total.Mercury.in.mg.g,.75)-quantile(Total.Mercury.in.mg.g,.25)
merc_outlier <- Total.Mercury.in.mg.g < quantile(Total.Mercury.in.mg.g,.25)-1.5*IQR_merc | Total.Mercury.in.mg.g > quantile(Total.Mercury.in.mg.g,.75)+1.5*IQR_merc
```

First, we check for outliers in the data-set:\
There are no outliners for number of meals, and there are `r sum(merc_outlier)` outliers for Total mercury level, and they are the following:`r Total.Mercury.in.mg.g[merc_outlier]`.\
However, base on the graph, and the fact that they are reasonably close to the upper limit for outliers (`r 1.5*IQR_merc+quantile(Total.Mercury.in.mg.g,.75)`), it's reasonable to include them in the data set for the following calculations.\
Base on the scatter plot, it's roughly a linear, positive relationship with moderate correlation between the number of meal with fish, and the total mercury level.\
Problem 2:\

```{r}
corXY_noOut <- round(cor(Number.of.meals.with.fish[!merc_outlier],Total.Mercury.in.mg.g[!merc_outlier]),4)
corXY <- round(cor(Number.of.meals.with.fish,Total.Mercury.in.mg.g),4)
```

The correlation is `r corXY`. Just to validate my judgement about outliers, the correlation without the outliers is `r corXY_noOut`. Both corelation values tell us that the relationship is positive with moderate strength.\
Problem 3:\

```{r}

lmFIT <- lm(Total.Mercury.in.mg.g ~ Number.of.meals.with.fish)
beta_1 <- lmFIT$coefficients[[2]]
beta_0 <- lmFIT$coefficients[[1]]
plot(Number.of.meals.with.fish, Total.Mercury.in.mg.g, xlab = "Number of Meal With Fish", 
     ylab = "Total Mercury Level(mg)", title("Plot of Numebr of meal with fish vs Total mercury level"), pch = 'x')
abline(beta_0,beta_1)

```

The least squares regression equation is $\hat y= \hat \beta_0*x+\hat \beta_1$, which for this data set is: y=`r beta_1`\*x+`r beta_0`\
Problem 4:\

```{r}
#can use the beta_0 and beta_1 value from the previous calculation as well.
beta_1_est <- round(corXY*merc_sd/meal_sd, 5)
beta_0_est <- round(merc_bar-beta_1*meal_bar,5)
```

The estimated $\hat \beta_0$ is `r beta_0_est`. It means that when there's no meal eaten with fish, the mercury level in the body is `r beta_0_est`mg.\
The estimated $\hat \beta_1$ is `r beta_1_est`. It meas that with our estimation based on the data set, the mercury level in the body increase linearly by `r beta_1_est` mg for every extra meal with fish eaten.\
Problem 5:\
Step 1:   $H_0: \beta_1=0$\
$H_1: \beta_1 \neq 0$ $\alpha = 0.05$ Step 2:\
$F = \frac{RegMS}{ResMS}$ For SLR, $F = ResDF*\frac{RegSS}{ResSS}$

```{r}
anova(lmFIT)
fcrit <- qf(.95,df1=1,df2=length(Total.Mercury.in.mg.g)-2)
```

Step 3:\
Reject $H_0$ if $F \geq F(1,n-2,\alpha)$\
$F(1,n-2,\alpha)$ = `r fcrit`\
Step 4: From the anova table, we have F = 93.689, greater than the fcrit Step 5:\
Reject $H_0$, at 95% confidence level, we have enough evidence that a significant linear relationship exist between the number of meals with fish and the total mercury level.

```{r}
Rsqr <- 309.24/(323.47+309.24)
```

The R square value is `r round(Rsqr,4)`, which means that `r round(Rsqr*100,4)`% of the variance in the total mercury level can be explained by the variance in the number of meals with fish.

```{r}
confLM <- confint(lmFIT,level = .9)
lowerCONF <- confLM[1]
upperCONF <- confLM[3]
```

The 90% confidence interval for $\beta_1$ is between `r lowerCONF` and `r upperCONF`.
