---
title: "HW5"
author: "Jiankun (Bob) Dong CM3226"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("pROC")
library(pROC)
# install.packages("caret")
library(caret)
library(car)
library(emmeans)
library(aod)
```

Probelm 1 and 2:

```{r }
dataSet <- read.csv("./A06.csv")
dataSet$temp_level <- as.numeric(dataSet$temp>=98.6)
dataSet$sex <- as.factor(dataSet$sex)
SexVTempLevel_T <- table(dataSet$temp_level,dataSet$sex)
colnames(SexVTempLevel_T) <- c('Male','Female')
SexVTempLevel_T
```

Problem 3: 1.\
$H_0 : p_1 = p_2$\
$H_1 : p_1 \neq p_2$\
$\alpha = 0.05$\
2. \
z = $\frac{\hat{p_1}-\hat{p_2}} {\hat{p}*(1-\hat{p})*(\frac{1}{n_1}+\frac{1}{n_2})}$ \
3. \
Decision rule: reject $H_0$ if \|z\|$\geq 1.96$ \
4. \

```{r}
prop.test(c(51,30),c(65,65),conf.level = 0.95, correct = TRUE)
p1_hat <- 51/65
p2_hat <- 30/65
p_hat <- 81/130
(z<-(p1_hat-p2_hat)/(p_hat*(1-p_hat)*(1/65+1/65)))
```

5.  Conclusion: reject $H_0$ since z is greater than 1.96. We reject the hypothesis that the proportion of people having high body temperature is the same across men and women.\

Problem 4:\

```{r}
m <- glm(dataSet$temp_level ~ dataSet$sex, family=binomial)

```

1.  $H_0 : \beta_1 = 0$\
    $H_1: \beta_1 \neq 0$\
    $\alpha = 0.05$\
2.  \
    z = $\frac {\beta_1} {SE\_{\beta\_1}}$ \
3.  \
    Decision rule: reject $H_0$ if \|z\| $\geq1.96$
4.  \

```{r}
summary(m)
```

z = 1.4469/0.3911 = `r 1.4469/0.3911` \> 1.96 5. Reject $H_0$ because z\>1.96. There is evidence of an association between sex and temperature level. The odds ratio for sex is `r exp(1.4469)` for change in sex. the associated 95% confidence interval is between `r exp(1.4469-1.96*0.3911)` and `r exp(1.4469+1.96*0.3911)`\
Problem 5:\

```{r}
dataSet$male <- ifelse(dataSet$sex == 1, 1, 0)
m2 <- glm(dataSet$temp_level ~ dataSet$male+dataSet$Heart.rate, 
         family=binomial)
summary(m2)

wald.test(b=coef(m2), Sigma=vcov(m2), Terms = 2:3)
```

Odds ratio for sex: `r exp(-1.38919)`\
Odds ratio for heart rate for 10 beat increase: `r exp(0.06337*10)`

```{r}
dataSet$prob_2 <- predict(m2, type=c("response"))
par(mfrow = c(1,1))
g2 <- roc(dataSet$temp_level ~ dataSet$prob_2)
auc(g2)
```

Problem6:\
The c statistic for the first model is:

```{r}
dataSet$prob_1 <- predict(m, type=c("response"))
par(mfrow = c(1,1))
g1 <- roc(dataSet$temp_level ~ dataSet$prob_1)
auc(g1)
```

And that's smaller than the second model. Therefore the second model is the better one with c statistic of 0.7289

```{r}
plot(1-g2$specificities, g2$sensitivities, 
     type="l", xlab="1-specificity", 
     ylab="Sensitivity", main="ROC curve")
abline(a=0, b=1)
grid()
```
